\section{Finding 2 — Pillar Discretisation and Interpolation Artefacts}
\label{sec:finding2}

\subsection{Finding statement}

The IR simulation model represents the yield curve on a finite set of maturity
pillars and relies on numerical interpolation to construct continuous curves.
This design introduces structural artefacts that may affect local curve smoothness,
implied forward rates, and exposure profiles. Model outputs may therefore exhibit
sensitivity to interpolation choices and maturity grid specification.

\subsection{Model features giving rise to the finding}

The finding arises from the following structural features:
\begin{itemize}
	\item Zero rates are simulated only at a discrete set of maturity pillars.
	\item Pricing and discounting require curve values at non-pillar maturities.
	\item Interpolation schemes are applied ex post and are not model-consistent.
\end{itemize}

As a result, numerical artefacts may be introduced independently of the underlying
stochastic dynamics.

\subsection{Risk implications for PFE}

Interpolation-induced artefacts may lead to:
\begin{itemize}
	\item Artificial oscillations in implied forward rates.
	\item Localised distortions of discount factors at intermediate maturities.
	\item Sensitivity of exposure profiles to curve discretisation choices.
\end{itemize}

These effects are particularly relevant for products sensitive to forward rates
or convexity, and for long-dated exposures.

\subsection{Validation strategy and diagnostics}

Three diagnostics are implemented to substantiate this finding:
\begin{enumerate}
	\item Interpolation sensitivity across numerical schemes.
	\item Roughness of implied forward-rate curves.
	\item Stability under changes in maturity pillar density.
\end{enumerate}

\subsection{Empirical evidence}

[Figures to be inserted here:
interpolation differences,
forward roughness plots,
pillar density comparison plots.]

\subsection{Conclusion}

The diagnostics demonstrate that the discrete pillar representation combined with
numerical interpolation introduces structural artefacts in the simulated term
structures. These artefacts are model-design driven and may materially affect
exposure calculations. Appropriate controls, monitoring, and sensitivity analyses
are therefore required.
